---
title: "my-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(StatComp19086)
```

## Preparation and Development Cycle

This is a vignette rmarkdown file. 
> Reference: https://r-pkgs.org/vignettes.html

### To create your first vignette, run:

`usethis::use_vignette("my-vignette")`
This will:

1. Create a vignettes/ directory.

2. Add the necessary dependencies to DESCRIPTION (i.e. it adds knitr to the `Suggests` and `VignetteBuilder` fields).

3. Draft a vignette, `vignettes/my-vignette.Rmd`. 

### Create a package bundle with the vignettes included

There are 2 methods:

1. `devtools::build_vignettes()` Build all vignettes from the console.

2. `devtools::build()` Create a package bundle with the vignettes included.

Note that 
```
#> devtools::build()
#>Building the package will delete...
#>  'C:/Users/Gnius/Documents/Code/R/StatComp-R/inst/doc'
#>Are you sure?
```

Note that
RStudio’s `Build & reload` does not build vignettes to save time. 
Similarly, `devtools::install_github()` (and friends) will not build vignettes by default because they’re time consuming and may require additional packages. 
You can force building with `devtools::install_github(build_vignettes = TRUE)`. 
This will also install all suggested packages.