---
title: "Homework 3"
author: "By 19086"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1111)
```

## Question 1 (6.5)
Suppose a $95\%$ symmetric t-interval is applied to estimate a mean, but the sample data are non-normal. 
Then the probability that the confidence interval covers the mean is not necessarily equal to $0.95$. 
Use a Monte Carlo experiment to estimate the coverage probability of the t-interval for random samples of $\chi^2(2)$ data with sample size $n = 20$. 
Compare your $t$-interval results with the simulation results in Example 6.4. 
(The t-interval should be more robust to departures from normality than the interval for variance.) 

## Question 2 (6.6)
Estimate the $0.025, 0.05, 0.95$, and $0.975$ quantiles of the skewness \sqrt{b_1} under normality by a Monte Carlo experiment. 
Compute the standard error of the estimates from (2.14) using the normal approximation for the density (with exact variance formula). 
Compare the estimated quantiles with the quantiles of the large sample approximation $\sqrt{b_1} \approx N(0, 6/n)$.

## Answer 1
Firstly, we use a Monte Carlo experiment to estimate the coverage probability of the t-interval for random samples of $\chi^2(2)$ data with sample size $n=20$. 

When we calculate the confidence intervals for one normally distributed population mean when population variance $\sigma^2$ is unknown, we use sample variance $s^2$ and obtain
\begin{equation}
  t = \frac{\bar{X} - \mu}{s/\sqrt{n}}
\end{equation}
which has the T-distribution with $n-1$ degrees of fredom, $t(n-1)$. 

A two side $100(1-\alpha)\%$ confidence interval is given by
\begin{equation}
  (\bar{X} - t_{\alpha/2} \cdot \frac{s}{\sqrt{n}}, \bar{X} + t_{\alpha/2} \cdot \frac{s}{\sqrt{n}})
\end{equation}
where $t_{\alpha/2}$ is the $\alpha/2$-quantile of the $t(n − 1)$ distribution.

```{r a3.1.1}
set.seed(111)
m <- 1000     # repeat
n <- 20       # sample size
alpha <- .05  # true confidence level

calcCI <- function(n, alpha) {
  y <- rchisq(n, df = 2)  # chi-square distribution
  return(c(mean(y),  sqrt(var(y)) * qt(alpha/2, df = n-1) / sqrt(n)))
}

UCL <- replicate(m, expr = calcCI(n = n, alpha = alpha))

X_bar <- UCL[1,]
X_bias <- UCL[2,]

# compute the mean to get the confidence level
num <- 0
for (ii in 1:m) {
  if ( (X_bar[ii]-X_bias[ii]) > 2 & (X_bar[ii]+X_bias[ii]) < 2)
    num <- num + 1
}
cat("Empirical confidence level: ", 100*num/m, "%\n")
```

Then we use a Monte Carlo experiment to estimate the coverage probability for random samples of $\chi^2(2)$ data with sample size $n=20$ using method in Example 6.4.

```{r a3.1.2}
set.seed(111)
m <- 1000     # repeat
n <- 20       # sample size
alpha <- .05  # true confidence level

calcCI <- function(n, alpha) {
  y <- rchisq(n, df = 2)  # chi-square distribution
  return((n-1) * var(y) / qchisq(alpha, df = n-1))
}

UCL <- replicate(m, expr = calcCI(n = n, alpha = alpha))

# compute the mean to get the confidence level
cat("Empirical confidence level: ", 100*mean(UCL > 4), "%\n")
```
Comparing my t-interval results with the simulation results in Example 6.4, we can find that the empirical confidence level of mine is much more precise than the result from Example 6.4. My result is close to the true level $1-alpha=95\%$. 


## Answer 2
The sample coefficient of skewness is denoted by $\sqrt{b_1}$,
\begin{equation}
  \sqrt{b_{1}} = 
    \frac{\frac{1}{n} \sum_{i=1}^{n}\left(X_{i}-\bar{X}\right)^{3}}
    {\left(\frac{1}{n}\sum_{i=1}^{n}\left(X_{i}-\bar{X}\right)^{2}\right)^{3 / 2}},
\end{equation}
which is asymptotically normal with mean $0$ and variance $6/n$ under normality assumption of $X$.

Firstly, we can calculate the quantiles of the the skewness as following
```{r a3.2.1}
# set.seed(111)
skewness <- function(x){
  x_bar <- mean(x)
  numerator <- mean((x-x_bar)^3)
  denominator <- (mean((x-x_bar)^2))^1.5
  return(numerator / denominator)
}
n <- 20
m <- 1000
mu <- 0
sigma <- 1
# generate samples
stat <- replicate(m, expr={
  x <- rnorm(n, mu, sigma)
  skewness(x)
})
estimator <- quantile(stat, c(.025, .05, .95, .975))
print(estimator)
```
The $0.025,0.05,0.95$, and $0.975$ quantiles of the skewness is shown above. 

Secondly, we compute the standard error of the estimates by equation (2.4), which shows that
\begin{equation}
  \operatorname{Var}\left(\hat{x}_{q}\right)=\frac{q(1-q)}{n f\left(x_{q}\right)^{2}}.
\end{equation}
If $f$ is the density of normal distribution, we can obtain the exact variance 
```{r a3.2.2}
level <- c(.025, .05, .95, .975)
se_xq <- sqrt(level*(1-level)/(n*(dnorm(level, mean=mu, sd=sigma)^2)))
print(round(rbind(level,se_xq), digits=5))
```

Finally,  we compare the estimated quantiles with the quantiles of the large sample approximation.
```{r a3.2.3}
xq <- qnorm(level, mean=0, sd=sqrt(6/n))
bound.upper <- estimator + 2*se_xq
bound.lower <- estimator - 2*se_xq
print(rbind(xq, estimator, bound.lower, bound.upper))
```

From the table, we can see that the estimators is approximate to the large sample approximation. The approximation $x_q$ is in the $2\sigma$ interval.

If we have more samples, for example $n=200$ rather than $n=20$, we obtain a more precise result as following
```{r a3.2.4, include=FALSE}
# set.seed(111)
skewness <- function(x){
  x_bar <- mean(x)
  numerator <- mean((x-x_bar)^3)
  denominator <- (mean((x-x_bar)^2))^1.5
  return(numerator / denominator)
}
n <- 200
m <- 1000
mu <- 0
sigma <- 1
# generate samples
stat <- replicate(m, expr={
  x <- rnorm(n, mu, sigma)
  skewness(x)
})
estimator <- quantile(stat, c(.025, .05, .95, .975))
# print(estimator)

level <- c(.025, .05, .95, .975)
se_xq <- sqrt(level*(1-level)/(n*(dnorm(level, mean=mu, sd=sigma)^2)))
# print(round(rbind(level,se_xq), digits=5))

xq <- qnorm(level, mean=0, sd=sqrt(6/n))
bound.upper <- estimator + 2*se_xq
bound.lower <- estimator - 2*se_xq
# print(rbind(xq, estimator, bound.lower, bound.upper))
```
```{r a3.2.5}
# n=200
print(rbind(xq, estimator, bound.lower, bound.upper))
```
