---
title: "Homework 8"
author: "By 19086"
date: "`r Sys.Date()`"
output: 
  html_document:
    # css: style.css

---

```{r setup, include=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(snowfall)
set.seed(1111)
```

## Question 1 (Exercises 11.1)

The natural logarithm and exponential functions are inverses of each other, so that mathematically $log(exp x) = exp(log x) = x$. 
Show by example that this property does not hold exactly in computer arithmetic. 
Does the identity hold with near equality? (See all.equal.)

## Solution 1
R provides the function *all.equal* to check for near equality of two *R* objects.
In a logical expression, use *isTRUE* to obtain a logical value. 

Firstly, we can code as following
```{r}
n <- 100
x <- seq(0.1, 2, length.out=n)
alpha <- unlist(base::lapply(x, function(x) ifelse(log(exp(x)) == exp(log(x)), 1, 0)))
cat("Ratio of equalities", sum(alpha)/n)
```

It means that there are about `r 100-sum(alpha)`\% $x$'s do not support $log(exp x) = exp(log x) = x$ in computer arithmetic. 

Secondly, 
```{r}
n <- 100
x <- seq(0.1, 2, length.out=n)
alpha <- unlist(base::lapply(x, function(x) ifelse(isTRUE(all.equal(log(exp(x)), exp(log(x)))), 1, 0)))
cat("Ratio of equalities", sum(alpha)/n)
```
We can see that the identity hold with near equality. 


## Question 2 (Exercises 11.5)
Write a function to solve the equation
\begin{equation}
  \frac{2 \Gamma\left(\frac{k}{2}\right)}{\sqrt{\pi(k-1)} \Gamma\left(\frac{k-1}{2}\right)} 
  \int_{0}^{c_{k-1}}\left(1+\frac{u^{2}}{k-1}\right)^{-k / 2} d u 
  = 
  \frac{2 \Gamma\left(\frac{k+1}{2}\right)}{\sqrt{\pi k} \Gamma\left(\frac{k}{2}\right)} 
  \int_{0}^{c_{k}}\left(1+\frac{u^{2}}{k}\right)^{-(k+1) / 2} d u 
\end{equation}
for $a$, where 
\begin{equation}
  c_{k}=\sqrt{\frac{a^{2} k}{k+1-a^{2}}}
\end{equation}
Compare the solutions with the points $A(k)$ in Exercise 11.4

### Exercise 11.4
Find the intersection points $A(k)$ in $(0, \sqrt{k})$ of the curves
\begin{equation}
  S_{k-1}(a) = P\left(t(k-1)>\sqrt{\frac{a^{2}(k-1)}{k-a^{2}}}\right)
\end{equation}
and
\begin{equation}
  S_{k}(a) = P\left(t(k)>\sqrt{\frac{a^{2} k}{k+1-a^{2}}}\right)
\end{equation}
for $k = 4 : 25, 100, 500, 1000$, where $t(k)$ is a Student $t$ random variable with $k$ degrees of freedom. 
(These intersection points determine the critical values for a $t$-test for scale-mixture errors proposed by Szekely [260].)



## Question 3 (A-B-O blood type problem)

Let the three alleles be A, B, and O with allele frequencies $p$, $q$, and $r$. 
The $6$ genotype frequencies under HWE and complete counts are as follows.

Genotype   | AA       |   BB     | OO       |  AO      | BO       | AB       |  Sum   |
-          | :-:      |  :-:     | :-:      | --:      | :-:      | :-:      | -:     |
Frequency  | $p^2$    | $q^2$    | $r^2$    | $2pr$    | $2qr$    | $2pq$    |  $1$   |
Count      | $n_{AA}$ | $n_{BB}$ | $n_{OO}$ | $n_{AO}$ | $n_{BO}$ | $n_{AB}$ |  $n$   |

Observed data: $n_{A\cdot} = n_{AA} + n_{AO} = 28$ (A-type), $n_{B·} = n_{BB} + n_{BO} = 24$ (B-type), $n_{OO} = 41$ (O-type), $n_{AB} = 70$ (AB-type).

1. Use EM algorithm to solve MLE of $p$ and $q$ (consider missing data $n_{AA}$ and $n_{BB}$). 

2. Show that the log-maximum likelihood values in M-steps are increasing via line plot.


## Answer 1

