---
title: "Homework 9"
author: "By 19086"
date: "`r Sys.Date()`"
output: 
  html_document:
    # css: style.css

---

```{r setup, include=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(snowfall)
set.seed(1111)
```

## Question 1 (Exercises 11.1)

The natural logarithm and exponential functions are inverses of each other, so that mathematically $log(exp x) = exp(log x) = x$. 
Show by example that this property does not hold exactly in computer arithmetic. 
Does the identity hold with near equality? (See all.equal.)

## Solution 1
R provides the function *all.equal* to check for near equality of two *R* objects.
In a logical expression, use *isTRUE* to obtain a logical value. 

Firstly, we can generate $100$ sample points from $[0.1, 2]$ uniformly, and check the identity. 
```{r}
n <- 100
x <- seq(0.1, 2, length.out=n)
alpha <- unlist(base::lapply(x, function(x) ifelse(log(exp(x)) == exp(log(x)), 1, 0)))
cat("Ratio of equalities", sum(alpha)/n)
```

It means that there are about `r 100-sum(alpha)`\% $x$'s do not support $log(exp x) = exp(log x) = x$ in computer arithmetic. 

Secondly, 
```{r}
n <- 100
x <- seq(0.1, 2, length.out=n)
alpha <- unlist(base::lapply(x, function(x) ifelse(isTRUE(all.equal(log(exp(x)), exp(log(x)))), 1, 0)))
cat("Ratio of equalities", sum(alpha)/n)
```
We can see that the identity hold with near equality. 


## Question 2 (Exercises 11.5)
Write a function to solve the equation
\begin{equation}
  \frac{2 \Gamma\left(\frac{k}{2}\right)}{\sqrt{\pi(k-1)} \Gamma\left(\frac{k-1}{2}\right)} 
  \int_{0}^{c_{k-1}}\left(1+\frac{u^{2}}{k-1}\right)^{-k / 2} d u 
  = 
  \frac{2 \Gamma\left(\frac{k+1}{2}\right)}{\sqrt{\pi k} \Gamma\left(\frac{k}{2}\right)} 
  \int_{0}^{c_{k}}\left(1+\frac{u^{2}}{k}\right)^{-(k+1) / 2} d u 
\end{equation}
for $a$, where 
\begin{equation}
  c_{k}=\sqrt{\frac{a^{2} k}{k+1-a^{2}}}
\end{equation}
Compare the solutions with the points $A(k)$ in Exercise 11.4

### Exercise 11.4
Find the intersection points $A(k)$ in $(0, \sqrt{k})$ of the curves
\begin{equation}
  S_{k-1}(a) = P\left(t(k-1)>\sqrt{\frac{a^{2}(k-1)}{k-a^{2}}}\right)
\end{equation}
and
\begin{equation}
  S_{k}(a) = P\left(t(k)>\sqrt{\frac{a^{2} k}{k+1-a^{2}}}\right)
\end{equation}
for $k = 4 : 25, 100, 500, 1000$, where $t(k)$ is a Student $t$ random variable with $k$ degrees of freedom. 
(These intersection points determine the critical values for a $t$-test for scale-mixture errors proposed by Szekely [260].)


## Solution 2

The equation contains integration calculation and ratio of large gamma function. 

I will use *integrate* function to calculate the integration as following
```{r, eval=FALSE}
# this is integrate function
myfun <- function(u, k){
  (1 + u^2/k)^(-(k+1) / 2)
}
integrate(myfun, lower=0, upper=ck, k=k)
```
And use *lgamma* to calculate the ratio of large gamma function
\begin{equation}
  \frac{\Gamma\left(\frac{k+1}{2}\right)}{\Gamma\left(\frac{k}{2}\right)}
\end{equation}
```{r, eval=FALSE}
exp(lgamma((k+1)/2) - lgamma(k/2))
```
To see how the result depends on $k$ and find the root of the equation, we can write the follwoing code using *uniroot* function.
```{r}
k.loop <- c(4:25, 100, 500, 1000)

myfun1 <- function(a, k.cur){
  ck_1 <- sqrt(a^2 * (k.cur-1) / (k.cur - a^2))
  ck <- sqrt(a^2 * k.cur / (k.cur + 1- a^2))
  fun.left <- 2 * exp(lgamma(k.cur/2) - lgamma((k.cur-1)/2)) / sqrt(pi * (k.cur-1)) * integrate(function(u, k) (1 + u^2/k)^(-(k+1) / 2), lower=0, upper=ck_1, k=k.cur-1)$value
  fun.right <- 2 * exp(lgamma((k.cur+1)/2) - lgamma(k.cur/2)) / sqrt(pi * k.cur) * integrate(function(u, k) (1 + u^2/k)^(-(k+1) / 2), lower=0, upper=ck, k=k.cur)$value
  fun.left - fun.right
}

# Record result 1
out1 <- numeric(length(k.loop))
ii <- 1
for (k in k.loop) {
  if (k>25){
    out1[ii] <- uniroot(myfun1, lower=0.01, upper=sqrt(k-1), k.cur=k)$root
  } else {
    out1[ii] <- uniroot(myfun1, lower=0.01*sqrt(k-1), upper=0.99*sqrt(k-1), k.cur=k)$root
  }
  
  ii <- ii + 1
}
```



Next, we compare the result with Exercise 11.4. 
```{r}
intersection <- function (k) {
  s.k.minus.one <- function (a) {
    1-pt(sqrt(a^2 * (k - 1) / (k - a^2)), df = k-1)
  } #the function of S_{k-1}(a)
  s.k <- function (a) {
    1-pt(sqrt(a^2 * k / (k + 1 - a^2)), df = k)
  } #the function of S_{k}(a)
  f <- function (a) {
    s.k(a) - s.k.minus.one(a)
  } #the root of f is the intersection points
  
  eps <- .Machine$double.eps^0.5 
  return(uniroot(f, interval = c(eps, sqrt(k)-eps))$root) #find the intersection points A(k) in (0,sqrt(k))
}

k <- c(4:25, 100, 500, 1000)
out2 <- sapply(k, function (k) {
  intersection(k)
  })
```

Print the result
```{r}
out <- cbind(k=k, Method1=out1, Method2=out2)
print(out)
```

We can find that method 1 is approximate to method 2 from Exercise 2 when $k\in \{4,\cdots, 25\}$. 
But when $k$ is larger, the root from method 1 is wrong. 
I check the value of $myfun1$, its value vibrates greatly rather than changes monotonically. This is due to the inaccurate calculation in *R*.


## Question 3 (A-B-O blood type problem)

Let the three alleles be A, B, and O with allele frequencies $p$, $q$, and $r$. 
The $6$ genotype frequencies under HWE and complete counts are as follows.

Genotype   | AA       |   BB     | OO       |  AO      | BO       | AB       |  Sum   |
-          | :-:      |  :-:     | :-:      | :--:     | :-:      | :-:      | -:     |
Frequency  | $p^2$    | $q^2$    | $r^2$    | $2pr$    | $2qr$    | $2pq$    |  $1$   |
Count      | $n_{AA}$ | $n_{BB}$ | $n_{OO}$ | $n_{AO}$ | $n_{BO}$ | $n_{AB}$ |  $n$   |

Observed data: $n_{A\cdot} = n_{AA} + n_{AO} = 28$ (A-type), $n_{B·} = n_{BB} + n_{BO} = 24$ (B-type), $n_{OO} = 41$ (O-type), $n_{AB} = 70$ (AB-type).

1. Use EM algorithm to solve MLE of $p$ and $q$ (consider missing data $n_{AA}$ and $n_{BB}$). 

2. Show that the log-maximum likelihood values in M-steps are increasing via line plot.


## Solution 3

### Idea of EM

The EM (Expectation Maximization) algorithm is a general optimization method that is often applied to find maximum likelihood estimates when data are incomplete.

1. Start with an initial estimate of the target parameter, and then alternate the E (expectation) step and M (maximization) step.

2. In the E step compute the conditional expectation of the objective function (usually a loglikelihood function) given the observed data and current parameter estimates.

3. In the M step, the conditional expectation is maximized with respect to the target parameter.

4. Update the estimates and iteratively repeat the E and M steps until the algorithm converges according to some criterion.

Because of the observed data, we can rewrite the table as:

| Genotype | Frequency   | Count             |
| -:       |:-:          | :-                |
|AA        |   $p^2$     |$n_{AA}$           |    
|BB        |   $q^2$     |$n_{BB}$           |    
|OO        |   $r^2$     |$n_{OO}=41$        |    
|AO        |   $2pr$     |$n_{AO}=28-n_{AA}$ |            
|BO        |   $2qr$     |$n_{BO}=24-n_{BB}$ |            
|AB        |   $2pq$     |$n_{AB}=70$        |    
|          |   $1$       |$n=163$            |

We can write code as following.

```{r}
# Write the log-likelihood function
lnL <- function(p, q, nA = 28, nB = 24, nOO = 41, nAB = 70) {
  r = 1 - p - q
  nA * log(p^2 + 2*p*r) + nB * log(q^2 + 2 * q * r) + 2 * nOO * log(r) + nAB * log(2 * p * q) 
}

# Write the E-M function
EM <- function (p, q, nA = 28, nB = 24, nOO = 41, nAB = 70, debug = FALSE) {
  
  # Evaluate the likelihood using initial estimates
  llk <- lnL(p, q, nA, nB, nOO, nAB)
  
  # Count the number of iterations so far
  iter <- 1
  # Vector for Record log-likelihood value
  llk.rec <- numeric(0)
  
  # Loop until convergence 
  while (TRUE)
  {
    # Estimate the frequency for allele O
    r = 1 - p - q
    
    # Record llk
    llk.rec <- c(llk.rec, llk)
    
    # First we carry out the E-step
    
    # The counts for genotypes O/O and A/B are effectively observed
    # Estimate the counts for the other genotypes
    nAA <- nA * p / (p + 2*r)
    nAO <- nA - nAA
    nBB <- nB * q / (q + 2*r)
    nBO <- nB - nBB
    
    # Print debugging information
    if (debug)
    {
      cat("Round #", iter, "lnLikelihood = ", llk, "\n")
      cat("    Allele frequencies: p = ", p, ", q = ", q, ", r = ", r, "\n")
      cat("    Genotype counts:    nAA = ", nAA, ", nAO = ", nAO, ", nBB = ", nBB, 
          ", nBO = ", nBO, "\n")
    }
    
    # Then the M-step
    p <- (2 * nAA + nAO + nAB) / (2 * (nA + nB + nOO + nAB))
    q <- (2 * nBB + nBO + nAB) / (2 * (nA + nB + nOO + nAB))
    
    # Then check for convergence 
    llk1 <- lnL(p, q, nA, nB, nOO, nAB)
    
    if (abs(llk1 - llk) < (abs(llk) + abs(llk1)) * 1e-6) 
      break       
    
    # Otherwise keep going
    llk <- llk1
    iter <- iter + 1
  }
  list(p = p, q = q, r=1-p-q, llk.rec=llk.rec)
}

# Set the initial estimate of the target parameters, then run the E-M function.
out <- EM(0.3, 0.25, nA = 28, nB = 24, nOO = 41, nAB = 70, debug = TRUE) 
cat("p=", out$p, "\tq=", out$q, "\tr=", out$r)
plot(1:length(out$llk.rec), out$llk.rec, "l", xlab="Iter", ylab="Log-likelihood")
```
